---
title: "class 5: Data Visulalization"
author: "Nikan"
format: pdf
---

# Plotting in R

R has multiple plotting and graphics systems. The most popular of which is **ggplot2**.

we have already played with "base" R graphics. this comes along with R "out of the box".

```{r}
plot(cars)
```

compared to base R plots ggplot is much more verbose - I need to write more code to get simple plots like the above.

have to first install ggplts

the install is a onetime only requirement. package is now on our computer. no need to reinstall. does have to be loaded into the project tho by calling `library()`.

```{r}
library(ggplot2)
```

```{r}
ggplot(cars)
```

all ggplot figures need at least 3 things: - data(this is the data.frame with our numbers ) - aesthetics ("aes", how our data mpas to the plot) - geomes (do want lines, points, columns, etc... )

```{r}
ggplot(data = cars) +
  aes(x=speed, y = dist) +
  geom_point()
```

i want a trend line to show the relationship between speed and stopping distance

```{r}
ggplot(data = cars) +
  aes(x=speed, y = dist) +
  geom_point() +
  geom_line()
```

thats not what we want

```{r}
ggplot(data = cars) +
  aes(x=speed, y = dist) +
  geom_point() + 
  geom_smooth()
```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
ggplot(genes) + 
  aes(x=Condition1, y=Condition2) +
  geom_point()
```

```{r}
p <- ggplot(genes) + 
  aes(x=Condition1, y=Condition2, col=State) +
  geom_point()
p
```

```{r}
p + scale_colour_manual( values=c("blue","gray","red") )
```

```{r}
p + scale_colour_manual(values=c("blue","gray","red")) +
    labs(title="Gene Expresion Changes Upon Drug Treatment",
         x="Control (no drug) ",
         y="Drug Treatment")
```










